---
title: "US Consumer Data"
description: |
  Summary: consumer sentiment improved with inflation tammed. However unemployment is uptrending. 
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{css}
.figure {
  text-align: center;
}
.figure p {
  font-size: 24px;
}
```

```{r, echo = FALSE, include=TRUE}

packages = c('readxl', 'tidyverse', 'ggridges', 'ggthemes', 'hrbrthemes', 'patchwork')

for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p) 
  }
  library(p, character.only = T)
}
```


```{r, echo = FALSE, include=TRUE}
uom_cs <- readxl::read_xlsx(path = "/Users/yapinghuang/Desktop/Year 2024/macro/data/data.xlsx", sheet = "UOM_CS")

cpi <- readxl::read_xlsx(path = "/Users/yapinghuang/Desktop/Year 2024/macro/data/data.xlsx", sheet = "CPI")

cpi_i <- readxl::read_xlsx(path = "/Users/yapinghuang/Desktop/Year 2024/macro/data/data.xlsx", sheet = "CPI_I")

unrate <- readxl::read_xlsx(path = "/Users/yapinghuang/Desktop/Year 2024/macro/data/data.xlsx", sheet = "unrate")

job_claim <- readxl::read_xlsx(path = "/Users/yapinghuang/Desktop/Year 2024/macro/data/data.xlsx", sheet = "Jb_claim")

retail <- readxl::read_xlsx(path = "/Users/yapinghuang/Desktop/Year 2024/macro/data/data.xlsx", sheet = "Retail")

```

1. UOM Consumer Sentiment

It rates the relative level of current and future economic conditions. 

Comments: consumer sentiment is far below pre-COVID, however, signifianct improvement observed in Jan 2024. 

```{r, echo = FALSE, fig.width=12, fig.height=8, include=TRUE}

p1 <- ggplot(data = uom_cs, 
       aes(x = Release_Date, y = Actual)) +
  geom_line()+
  theme_economist(
              base_size = 20)+
  labs(x = "Year",
       y = "Consumer sentiment",
       subtitle = "A survey of around 500 consumers",
       caption = "Source: University of Michigan")+
  ggtitle("Consumer Sentiment Still NOT Recovered to pre-COVID level")

start_date <- as.POSIXct("2023-7-13")
last_6M_data <- uom_cs[uom_cs$Release_Date >= start_date,]

p2 <- ggplot(data = last_6M_data, 
       aes(x = Release_Date, y = Actual)) +
  geom_line(color = "green", size = 1.5)+
  geom_vline(aes(xintercept=as.POSIXct("2023-11-30")),
             color="red",
             linetype="dashed",size=1) +
  theme_ipsum(axis_title_size = 12,
              base_size = 8,
              grid = "Y")+
  labs(x = "",
       y = "")+
  ggtitle("CS improved in Jan 2024")

p1 + inset_element(p2, 
                   left = 0.65, 
                   bottom = 0.5, 
                   right = 1, 
                   top = 1)

```

2. Consumer Price Index (CPI)

Consumer Price Index (CPI) measures the monthly change in prices paid by U.S. consumers. 
The Bureau of Labor Statistics (BLS) calculates the CPI as a weighted average of prices for a basket of goods and services representative of aggregate U.S. consumer spending.
Index 1982-1984=100, Seasonally Adjusted. 

Comments: CPI has dropped from highest point at 9% to 3.3% in Dec 2023; however it is still higher than FED target at 2%. 

```{r, echo = FALSE, fig.width=12, fig.height=8, include=TRUE}

start_date <- as.POSIXct("2017-01-01")
cpi <- cpi[cpi$Date >= start_date,]
cpi_i <- cpi_i[cpi_i$Date >= start_date,]
coeff <- 200

max_point <- cpi_i[which.max(cpi_i$CPI_I), ]

p1 <- ggplot(data = cpi, 
       aes(x = Date, y=CPI)) +
  geom_line()+
  theme_economist(
              base_size = 20)+
  labs(x = "Year",
       y = "CPI",
       title = "CPI for All Urban Consumers: All Items Less Food and Energy in U.S. City Average",
       caption = "Source: FRED ECONOMIC DATA")


p2 <- ggplot(data = cpi_i, 
       aes(x = Date, y = CPI_I)) +
  geom_line(color = "brown")+
  theme_economist(
              base_size = 20)+
  labs(x = "Year",
       y = "CPI Rate",
       title = "CPI: All Items: Total for United States",
       caption = "Source: OECD Main Economic Indicators - complete database")+
    geom_text(data = max_point, aes(label = paste("Max: ", round(CPI_I, digits = 2))), vjust = -0.2, hjust = 0.8)

p1/p2

```

3. Unemployment Rate (Percent)

The unemployment rate represents the number of unemployed as a percentage of the labor force. Labor force data are restricted to people 16 years of age and older, who currently reside in 1 of the 50 states or the District of Columbia, who do not reside in institutions (e.g., penal and mental facilities, homes for the aged), and who are not on active duty in the Armed Forces.

This rate is also defined as the U-3 measure of labor underutilization.

Comments: Unemployment Rate largely recovered to pre-COVID level, however, recent 12 month data is trending upward. 

```{r, echo = FALSE, fig.width=12, fig.height=8, include=TRUE}

start_date <- as.POSIXct("2017-01-01")
start_date_1year <- as.POSIXct("2022-12-01")

unrate <- unrate[unrate$date >= start_date,]
unrate_1year <- unrate[unrate$date >= start_date_1year,]

max_point <- unrate[which.max(unrate$UNRATE), ]

p1 <- ggplot(data = unrate, 
       aes(x = date, y = UNRATE)) +
  geom_line()+
  theme_economist()+
  labs(x = "Year",
       y = "Unemployment rate in %",
       title = "Personal Consumption Expenditures",
       caption = "Source: U.S. Bureau of Labor Statics")+
  geom_text(data = max_point, aes(label = paste("Max: ", UNRATE)), vjust = -0.2, hjust = 0.8)


p2 <- ggplot(data = unrate_1year, 
       aes(x = date, y = UNRATE)) +
  geom_line()+
  theme_economist()+
  labs(x = "",
       y = "",
       title = "Past 1 year unemployment rate")

p1 + inset_element(p2, 
                   left = 0.55, 
                   bottom = 0.5, 
                   right = 1, 
                   top = 0.95)

```
4. Initial and Continued Claims (CCSA, ICSA)

An initial claim is a claim filed by an unemployed individual after a separation from an employer. The claim requests a determination of basic eligibility for the Unemployment Insurance program.

Continued claims, also referred to as insured unemployment, is the number of people who have already filed an initial claim and who have experienced a week of unemployment and then filed a continued claim to claim benefits for that week of unemployment. Continued claims data are based on the week of unemployment, not the week when the initial claim was filed.

```{r echo=FALSE, fig.width=12, fig.height=8, include=TRUE}

start_date <- as.POSIXct("2019-01-01")
end_date <- as.POSIXct("2023-12-09")
job_claim <- job_claim [job_claim$date >= start_date & job_claim$date <= end_date,]
job_claim$CCSA_in_millions <- job_claim$CCSA / 1e6
job_claim$ICSA_in_millions <- job_claim$ICSA / 1e6

ggplot(data = job_claim, 
       aes(x = date)) +
  geom_line(aes(y = CCSA_in_millions), color = "red")+
  geom_line(aes(y = ICSA_in_millions), color = "black", linetype = "dashed")+
  theme_economist()+
  labs(x = "Year",
       y = "CCSA/ICSA in millions",
       title = "Weekly Initial and Continued Claims",
       caption = "Source: U.S. Employment and Training Administration")
```


5. Retails Sales 

E-commerce sales are included. 

```{r, echo = FALSE, fig.width=12, fig.height=8, include=TRUE}
start_date <- as.POSIXct("2019-01-01")
end_date <- as.POSIXct("2023-12-01")
retail <- retail[retail$date >= start_date & retail$date <= end_date,]

ggplot(data = retail, 
       aes(x = date, y = RSXFS)) +
  geom_line()+
  theme_economist()+
  labs(x = "Year",
       y = "Retail Sales",
       title = "Retails Sales",
       caption = "Source: U.S. Census Bureau")
```