---
title: "US Consumer Data"
description: |
  ALL consumer data in one place
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{css}
.figure {
  text-align: center;
}
.figure p {
  font-size: 24px;
}
```

```{r, echo = FALSE, include=TRUE}

packages = c('readxl', 'tidyverse', 'ggridges', 'ggthemes', 'hrbrthemes', 'patchwork')

for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p) 
  }
  library(p, character.only = T)
}
```


```{r, echo = FALSE, include=TRUE}
uom_cs <- readxl::read_xlsx(path = "/Users/yapinghuang/Desktop/Year 2024/macro/data/data.xlsx", sheet = "UOM_CS")

cpi <- readxl::read_xlsx(path = "/Users/yapinghuang/Desktop/Year 2024/macro/data/data.xlsx", sheet = "CPI")

cpi_i <- readxl::read_xlsx(path = "/Users/yapinghuang/Desktop/Year 2024/macro/data/data.xlsx", sheet = "CPI_I")

cpe <- readxl::read_xlsx(path = "/Users/yapinghuang/Desktop/Year 2024/macro/data/data.xlsx", sheet = "CPE")

job_claim <- readxl::read_xlsx(path = "/Users/yapinghuang/Desktop/Year 2024/macro/data/data.xlsx", sheet = "Jb_claim")

retail <- readxl::read_xlsx(path = "/Users/yapinghuang/Desktop/Year 2024/macro/data/data.xlsx", sheet = "Retail")

```

1. UOM Consumer Sentiment

It rates the relative level of current and future economic conditions. 

Comments: consumer sentiment is far below pre-COVID, however, signifianct improvement observed in Jan 2024. 

```{r, echo = FALSE, fig.width=12, fig.height=8, include=TRUE}

p1 <- ggplot(data = uom_cs, 
       aes(x = Release_Date, y = Actual)) +
  geom_line()+
  theme_economist(
              base_size = 20)+
  labs(x = "Year",
       y = "Consumer sentiment",
       subtitle = "A survey of around 500 consumers",
       caption = "Source: University of Michigan")+
  ggtitle("Consumer Sentiment Still NOT Recovered to pre-COVID level")

start_date <- as.POSIXct("2023-7-13")
last_6M_data <- uom_cs[uom_cs$Release_Date >= start_date,]

p2 <- ggplot(data = last_6M_data, 
       aes(x = Release_Date, y = Actual)) +
  geom_line(color = "green", size = 1.5)+
  geom_vline(aes(xintercept=as.POSIXct("2023-11-30")),
             color="red",
             linetype="dashed",size=1) +
  theme_ipsum(axis_title_size = 12,
              base_size = 8,
              grid = "Y")+
  labs(x = "",
       y = "")+
  ggtitle("CS improved in Jan 2024")

p1 + inset_element(p2, 
                   left = 0.65, 
                   bottom = 0.5, 
                   right = 1, 
                   top = 1)

```

2. Consumer Price Index (CPI)

Consumer Price Index (CPI) measures the monthly change in prices paid by U.S. consumers. 
The Bureau of Labor Statistics (BLS) calculates the CPI as a weighted average of prices for a basket of goods and services representative of aggregate U.S. consumer spending.
Index 1982-1984=100, Seasonally Adjusted. 

Inflation Rate= (New CPI−Prior CPI)/Prior CPI x 100

Comments: CPI has dropped from highest point at 9% to 3.3% in Dec 2023; however it is still higher than FED target at 2%. 

```{r, echo = FALSE, fig.width=12, fig.height=8, include=TRUE}

start_date <- as.POSIXct("2017-01-01")
cpi <- cpi[cpi$Date >= start_date,]
cpi_i <- cpi_i[cpi_i$Date >= start_date,]
coeff <- 200

max_point <- cpi_i[which.max(cpi_i$CPI_I), ]

p1 <- ggplot(data = cpi, 
       aes(x = Date, y=CPI)) +
  geom_line()+
  theme_economist(
              base_size = 20)+
  labs(x = "Year",
       y = "CPI",
       title = "CPI for All Urban Consumers: All Items Less Food and Energy in U.S. City Average",
       caption = "Source: FRED ECONOMIC DATA")


p2 <- ggplot(data = cpi_i, 
       aes(x = Date, y = CPI_I)) +
  geom_line(color = "brown")+
  theme_economist(
              base_size = 20)+
  labs(x = "Year",
       y = "CPI Rate",
       title = "CPI: All Items: Total for United States",
       caption = "Source: OECD")+
    geom_text(data = max_point, aes(label = paste("Max: ", CPI_I)), vjust = -0.2, hjust = 0.8)

p1/p2

```

3. Personal Consumption Expenditures Price Index (PCE)

In Billions of Dollars

```{r, echo = FALSE, fig.width=12, fig.height=8, include=TRUE}

start_date <- as.POSIXct("2019-01-01")
end_date <- as.POSIXct("2023-11-01")
cpe <- cpe[cpe$observation_date >= start_date & cpe$observation_date <= end_date,]

ggplot(data = cpe, 
       aes(x = observation_date, y = PCE)) +
  geom_line()+
  theme_economist()+
  labs(x = "Year",
       y = "CPE in US$bn",
       title = "Personal Consumption Expenditures",
       caption = "Source: U.S. Bureau of Economic Analysis")
```
4. Initial and Continued Claims (CCSA, ICSA)

Source: U.S. Employment and Training Administration 

An initial claim is a claim filed by an unemployed individual after a separation from an employer. The claim requests a determination of basic eligibility for the Unemployment Insurance program.

Continued claims, also referred to as insured unemployment, is the number of people who have already filed an initial claim and who have experienced a week of unemployment and then filed a continued claim to claim benefits for that week of unemployment. Continued claims data are based on the week of unemployment, not the week when the initial claim was filed.

```{r echo=FALSE, fig.width=12, fig.height=8, include=TRUE}

start_date <- as.POSIXct("2019-01-01")
end_date <- as.POSIXct("2023-12-09")
job_claim <- job_claim [job_claim$date >= start_date & job_claim$date <= end_date,]
job_claim$CCSA_in_millions <- job_claim$CCSA / 1e6
job_claim$ICSA_in_millions <- job_claim$ICSA / 1e6

ggplot(data = job_claim, 
       aes(x = date)) +
  geom_line(aes(y = CCSA_in_millions), color = "red")+
  geom_line(aes(y = ICSA_in_millions), color = "black", linetype = "dashed")+
  theme_economist()+
  labs(x = "Year",
       y = "CCSA/ICSA in millions",
       title = "Weekly Initial and Continued Claims")
```


5. Retails Sales 

Source: U.S. Census Bureau  

E-commerce sales are included in the total monthly sales estimates.
The value for the most recent month is an advance estimate that is based on data from a subsample of firms from the larger Monthly Retail Trade Survey. The advance estimate will be superseded in following months by revised estimates derived from the larger Monthly Retail Trade Survey.

```{r, echo = FALSE, fig.width=12, fig.height=8, include=TRUE}
start_date <- as.POSIXct("2019-01-01")
end_date <- as.POSIXct("2023-12-01")
retail <- retail[retail$date >= start_date & retail$date <= end_date,]

ggplot(data = retail, 
       aes(x = date, y = RSXFS)) +
  geom_line()+
  theme_economist()+
  labs(x = "Year",
       y = "Retail Sales",
       title = "Retails Sales")
```