---
title: "US Consumer Data"
description: |
  ALL consumer data in one place
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{css}
.figure {
  text-align: center;
}
.figure p {
  font-size: 24px;
}
```

```{r, echo = FALSE, include=TRUE}

packages = c('readxl', 'tidyverse', 'ggridges', 'ggthemes')

for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p) 
  }
  library(p, character.only = T)
}
```


```{r, echo = FALSE, include=TRUE}
uom_cs <- readxl::read_xlsx(path = "/Users/yapinghuang/Desktop/Year 2024/macro/data/data.xlsx", sheet = "UOM_CS")

cpi <- readxl::read_xlsx(path = "/Users/yapinghuang/Desktop/Year 2024/macro/data/data.xlsx", sheet = "CPI(2)")

cpe <- readxl::read_xlsx(path = "/Users/yapinghuang/Desktop/Year 2024/macro/data/data.xlsx", sheet = "CPE")

job_claim <- readxl::read_xlsx(path = "/Users/yapinghuang/Desktop/Year 2024/macro/data/data.xlsx", sheet = "Jb_claim")

retail <- readxl::read_xlsx(path = "/Users/yapinghuang/Desktop/Year 2024/macro/data/data.xlsx", sheet = "Retail")

```

1. UOM Consumer Sentiment

Source: University of Michigan

```{r, echo = FALSE, fig.width=16, fig.height=12, include=TRUE}

ggplot(data = uom_cs, 
       aes(x = Release_Date, y = Actual)) +
  geom_line()+
  theme_economist()+
  labs(x = "Year",
       y = "Consumer Sentiment",
       title = "UOM Consumer Sentiment")

```

2. Consumer Price Index (CPI)

Source: U.S. Bureau of Labor Statistics  

Consumer Price Index for All Urban Consumers: All Items in U.S. City Average

```{r, echo = FALSE, fig.width=16, fig.height=12, include=TRUE}

start_date <- as.POSIXct("2019-01-01")
cpi <- cpi[cpi$Date >= start_date,]

ggplot(data = cpi, 
       aes(x = Date, y = CPI)) +
  geom_line()+
  theme_economist()+
  labs(x = "Year",
       y = "CPI",
       title = "CPI since year 2019")

```

3. Personal Consumption Expenditures Price Index (PCE)

Source: U.S. Bureau of Economic Analysis  
In Billions of Dollars

```{r, echo = FALSE, fig.width=16, fig.height=12, include=TRUE}

start_date <- as.POSIXct("2019-01-01")
end_date <- as.POSIXct("2023-11-01")
cpe <- cpe[cpe$observation_date >= start_date & cpe$observation_date <= end_date,]

ggplot(data = cpe, 
       aes(x = observation_date, y = PCE)) +
  geom_line()+
  theme_economist()+
  labs(x = "Year",
       y = "CPE in US$bn",
       title = "Personal Consumption Expenditures")
```


4. Initial and Continued Claims (CCSA, ICSA)

Source: U.S. Employment and Training Administration 

An initial claim is a claim filed by an unemployed individual after a separation from an employer. The claim requests a determination of basic eligibility for the Unemployment Insurance program.

Continued claims, also referred to as insured unemployment, is the number of people who have already filed an initial claim and who have experienced a week of unemployment and then filed a continued claim to claim benefits for that week of unemployment. Continued claims data are based on the week of unemployment, not the week when the initial claim was filed.

```{r echo=FALSE, fig.width=16, fig.height=12, include=TRUE}

start_date <- as.POSIXct("2019-01-01")
end_date <- as.POSIXct("2023-12-09")
job_claim <- job_claim [job_claim$date >= start_date & job_claim$date <= end_date,]
job_claim$CCSA_in_millions <- job_claim$CCSA / 1e6
job_claim$ICSA_in_millions <- job_claim$ICSA / 1e6

ggplot(data = job_claim, 
       aes(x = date)) +
  geom_line(aes(y = CCSA_in_millions), color = "red")+
  geom_line(aes(y = ICSA_in_millions), color = "black", linetype = "dashed")+
  theme_economist()+
  labs(x = "Year",
       y = "CCSA/ICSA in millions",
       title = "Weekly Initial and Continued Claims")
```


5. Retails Sales 

Source: U.S. Census Bureau  

E-commerce sales are included in the total monthly sales estimates.
The value for the most recent month is an advance estimate that is based on data from a subsample of firms from the larger Monthly Retail Trade Survey. The advance estimate will be superseded in following months by revised estimates derived from the larger Monthly Retail Trade Survey.

```{r, echo = FALSE, fig.width=16, fig.height=12, include=TRUE}
start_date <- as.POSIXct("2019-01-01")
end_date <- as.POSIXct("2023-12-01")
retail <- retail[retail$date >= start_date & retail$date <= end_date,]

ggplot(data = retail, 
       aes(x = date, y = RSXFS)) +
  geom_line()+
  theme_economist()+
  labs(x = "Year",
       y = "Retail Sales",
       title = "Retails Sales")
```